---
title: "day6-wrangling-tidyverse"
output: html_document
date: '2022-08-15'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}


library(tidyverse)
library(tidyr)
library(dplyr)
library(here)
library(janitor)

```

## Read in the WB data files

```{r}

wb_indicators <- read_csv(here::here("data", "wb_indicators.csv"), na = c("..", ""))
wb_metadata <- read_csv(here::here("data", "wb_indicators_metadata.csv"))

```

## Tidy up my data

### Using pivot_longer to get years into a single column

```{r}

wb_indicators_long <- wb_indicators %>% 
  pivot_longer(cols = '2001 [YR2001]':'2020 [YR2020]',
               names_to = "year",
               values_to = "indicator_value")
```

### seperate to clean up the year column

```{r}


wb_clean <- wb_indicators_long %>%
  separate(col = year, into = c("year", "year_chr"), sep = " ") %>% # Separates
  select(-year_chr, -"Country Code", -"Series Code") %>% # Also exclude Country Code and Series Code
  mutate(year = as.numeric(year)) %>% # Makes the year column into numeric
  drop_na('Series Name') %>% # Drops all (entire) rows that have NA values
  pivot_wider(names_from = 'Series Name',
              values_from = indicator_value)

names(wb_clean) <- c("country", "year", "access_clean_fuels_pp", "access_electricity_pp", "co2_emmisions_kt", "fossil_fuel_consumption_pct", "water_stress")


# Widen the variables currently trapped in 'Series Name' to be spread across multiple columns, and put the values that populated the columns from 'indicator_value'

wb_subset <- wb_clean %>%
  filter(country %in% c("Algeria", "Barbados", "Argentina", "France", "Chile"))

```

## Plotting a graph

```{r}

# ggplot(data = wb_subset, aes(x = year, co2_emmisions_kt, group = country)) +
# geom_line()

# SAME THING

ggplot(data = wb_subset, aes(x = year, co2_emmisions_kt)) +
  geom_line(aes(color = country)) +
  facet_wrap(~country, ncol = 5)


```

## Filtering by rows

Example 1: Filter to only keep exact matches

```{r}
ex_1 <- starwars %>% 
  filter(mass > 180)
# Check code with min(ex_1$height)
```

```{r}
ex_2 <- starwars %>% 
  filter(eye_color == "blue")
# Check code with unique(ex_2$eye_color)
```

# Subset that only contains homeworld = Naboo

```{r}
ex_3 <- starwars %>% 
  filter(homeworld == "Naboo")
```

# Filter for 2 things to be true

```{r}
ex_4 <- starwars %>% 
  filter(height > 180 & homeworld == "Tatooine")
```

# Subset where hair color is brown & species is human

```{r}
ex_5 <- starwars %>%
  filter(hair_color == "brown" & species == "Human")
```

### Filtering OR statements

```{r}
ex_6 <- starwars %>% 
  filter(height > 180 | eye_color == "yellow")

ex_61 <- starwars %>%
  filter(homeworld == "NA")
```

# Subset that keeps rows if homeworld is Endor OR species is droid

```{r}
ex_7 <- starwars %>% 
  filter(homeworld == "Endor" | species == "Droid")

ex_71 <- starwars %>% 
  filter(homeworld == "Endor")
```

# Keep observations where eye color are blue, brown red

```{r}
ex_8 <- starwars %>%
  filter(eye_color == "blue" | eye_color == "brown" | eye_color == "red")

# OR

ex_9 <- starwars %>% 
  filter(eye_color %in% c("blue", "brown", "red"))
# goes through by row and asks does this value equal blue, brown or red?


## THIS IS NEVER WHAT YOU WANT TO DO
ex_10 <- starwars %>%
  filter(eye_color == c("blue", "brown", "red"))
# would go through and ask does eye_color value match blue, then goes to next row and asks does eye_color match brown, then goes to 3rd row and asks does eye_color match red, 4th row blue


# create a subset where homeworld is Endor OR Tatooine OR Naboo AND height < 150

ex_11 <- starwars %>% 
  filter((homeworld %in% c("Endor", "Tatooine", "Naboo")),
         height < 150)
```

# Examples with negate !

```{r}
ex_12 <- starwars %>% 
  filter(species != "Human")

ex_13 <- starwars %>%
  filter(!species %in% c("Human", "Ewok", "Wookie"))
```

## `select()`: select or exclude columns

```{r}
# Select by name
ex_14 <- wb_clean %>% 
  select(country, year, co2_emmisions_kt)

# Select a range of columns

ex_15 <- wb_clean %>% 
  select(year:fossil_fuel_consumption_pct)

#Select a range of columns excluding columns within range

ex_16 <- wb_clean %>% 
  select(year:fossil_fuel_consumption_pct, -access_electricity_pp)

```

```{r}
ex_17 <- wb_clean %>% 
  filter(country %in% c("Bolivia", "Chile", "Mexico")) %>%
  select(country:fossil_fuel_consumption_pct)
```

# Select certain columns and renames them in order

```{r}
ex_18 <- wb_clean %>% 
  select(year_new = year,
         emissions = co2_emmisions_kt,
         country_name = country)

# Select only columns for country, year and water stress, and reorder them as year, country and water stress, then filter to only include observations for any three countries of my choosing

ex_19 <- wb_clean %>% 
  select(year,
         country, 
         water_stress) %>%
  filter(country %in% c("Argentina", "France", "Kenya"))
```

### `rename` rename columns

```{r}
ex_20 <- wb_clean %>% 
  rename(ffcp = fossil_fuel_consumption_pct,
         ws = water_stress)

# Rename any 2 columns

ex_21 <- wb_clean %>% 
  rename(access_elec_pp = access_electricity_pp,
         co2_emmis_kt = co2_emmisions_kt)
```

### `mutate()`: update an existing column, or add a new one

If you assign a NEW column name, it will add a new column to the end of your data frame. If you assign an existing column name, it will overwrite that column

```{r}
ex_22 <- wb_clean %>% 
  mutate(co2_emissions_t = co2_emmisions_kt * 1000) %>% 
  relocate(co2_emissions_t, .after = co2_emmisions_kt)

ex_23 <- wb_clean %>% 
  mutate(yay = "YAY")

ex_24 <-wb_clean %>% 
  mutate(year = as.character(year))

#dplyr::relocate()

ex_26 <- wb_clean %>% 
  relocate(water_stress, .after = year)

# Relocate co2 emissions in tons to be next to emissions in kilotons

```

### `group_by %>% summarize`

group_by: Creates groups under the hood based on variables you specify

summarize: apply your functions to EACH GROUP, and return them in a nice summary table

```{r}
ex_27 <- starwars %>% 
  filter(homeworld %in% c("Naboo", "Tatooine")) %>% 
  group_by(species) %>% 
  summarise(mean_height = mean(height, na.rm = TRUE),
            mean_mass = mean(mass, na.rm = TRUE),
            count = n())
```

```{r}

df <- tribble(
  ~species, ~length, ~number,
  "lobster", 12, 4,
  "lobster", 14, 6,
  "lobster", 15, 2
)

df_case <- df %>% 
  uncount(number)
```

```{r}
# count counts observations
# combination of group_by, summarize, n()
starwars %>% 
  count(homeworld, species)
# SAME AS
starwars %>% 
  group_by(homeworld, species) %>% 
  summarize(size = n())
```

Practice Example:

- Start from wb_clean
- Only keep observations for Nicaragua, Ecuador, Peru, and Chile, AND for years greater than 2005
- Only keep columns for country, year and CO2 emissions
- Add a new column that is co2 emissions converted to megatons
- find the total co2 emissions for the years still in the data set by country

```{r}
final_ex <- wb_clean %>% 
  filter(country %in% c("Nicaragua", "Ecuador", "Peru", "Chile"),
                        year > 2005) %>% 
  select(country, year, co2_emmisions_kt) %>% 
  mutate(co2_emissions_mt = co2_emmisions_kt / 1000) %>% 
  group_by(country) %>% 
  summarise(sum_co2_mt = sum(co2_emissions_mt, na.rm = TRUE))
```

